---
title: "ActivityMonitorData"
author: "Reynold Guerrier"
date: "Sunday, June 14, 2015"
output: html_document
---

#Peer Assessment 1
=====================================================================================================================
This assignment makes use of data from a personnal activity monitoring device. The devices collects data at 5 minutes interval through out the day. The data consists of two months of data of an anonymous individual collected during the months of October & November 2012 and include a number of steps taken in 5 minute intervals eah day

###Data Processing
The dataset: Activity Monitoring Data used for this assignment has been downloaded (https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip)

1. Download the Activity file in the current working directory
2. Uncompress it and convert it as a dataframe 
3. Define the class of each column 
```{r, echo=TRUE}
library(ggplot2)
library(plyr)
setwd("c:\\data\\perso\\cours-online\\coursera\\reprodresearch\\data")
mondata <- read.csv("activity.csv", colClasses = c("numeric", "Date","numeric"))
PerDay <- aggregate(steps ~ date, mondata, sum, na.action = na.pass)
PerDay <- cbind(PerDay, label = rep("with.na", nrow(PerDay)))
ggplot(PerDay, aes(x = steps)) + geom_histogram(binwidth =1500, colour = "black", fill = "green") + labs(title = "Number of mean steps taken daily", x = "Number of steps", y = "Frequency")
```




The number of steps taken per Day are:

* Mean:   10766.19
* Median: 10765

### What is the average daily activity pattern


```{r, echo=TRUE}
perInterval <- aggregate(steps ~ interval, mondata, mean, na.rm = TRUE)
ggplot(perInterval, aes(x = interval, y = steps)) + geom_line() + labs(title = "Average of steps taken daily", x = "Interval", y = "Number of Steps")
```



### Imputing missing values

```{r, echo=TRUE}
mondata.impute <- adply(mondata, 1, function(x) if (is.na(x$steps)) {x$steps = round(perInterval[perInterval$interval == x$interval,2]) 
                                                                     x}
                        else {
                          x
                        })

```

We will obtain the following histogram

```{r, echo=TRUE}
# We add label for tracking purpose
perDay.impute <- aggregate(steps ~ date, mondata.impute, sum)
perDay.impute <- cbind(perDay.impute, label = rep("without.na", nrow(perDay.impute)))
ggplot(perDay.impute, aes(x = steps)) + geom_histogram(binwidth = 1500, colour = "black", fill = "green") + labs(title = "Steps taken per Day", x = "Number of Steps", y = "Frequency") + theme(legend.position = "bottom")


```

We observe that he mean value and the median value have shifted a little

* Mean:     10765.64
* Median:   10762

Below we have the 2 histograms


```{r, echo=TRUE}
perDay.all <- rbind(PerDay, perDay.impute)
levels(perDay.all$label) <- c("With NA", "Without NA")
ggplot(perDay.all, aes(x = steps, fill = label)) + geom_histogram(binwith = 1500, colour = "black", alpha = 0.2) + labs(title = "Steps taken per Day", x = "Number of Steps", y = "Frequency") + theme(legend.position = "bottom")

```

## Are there differences in activity patterns between weekdays and weekends?

To do this comparison with the table with filled-in missing values, we follow the next steps:

 1. I subset the data in 2 parts One for Weekdays and another fot=r Weekend
2. I calculate the average steps for each dataset
3. label them and 
4.  plot the two datasets for comparison.

```{r, echo=FALSE}
#Set the system to the local time
Sys.setlocale(locale = "C")

```

```{r, echo=TRUE}
mondata.weekend <- subset(mondata.impute, weekdays(date) %in% c("Saturday", "Sunday"))
mondata.weekday <- subset(mondata.impute, !weekdays(date) %in% c("Saturday", "Sunday"))
mondata.weekend <- aggregate(steps ~ interval, mondata.weekend, mean)
mondata.weekday <- aggregate(steps ~ interval, mondata.weekday, mean)
mondata.weekend <- cbind(mondata.weekend, day = rep("weekend"))
mondata.weekday <- cbind(mondata.weekday, day = rep("weekday"))
# Combine the 2 subsets 
mondata.week <- rbind(mondata.weekend, mondata.weekday)
levels(mondata.week$day) <- c("Weekend", "Weekday")
ggplot(mondata.week, aes(x = interval, y = steps)) + geom_line() + facet_grid(day ~ 
    .) + labs(x = "Interval", y = "Number of steps")


```
